---
title: "NBS_estimate"
output: 
  html_document
date: "`r Sys.Date()`"
---


## Prepare data  
  
```{r setup, eval = FALSE, include=FALSE}
### This section repeats the process of mapping NBS approaches, using population, land cover and emission by sector data 
### Skip it if you already have the color-coded maps** - Jump to the next section to bring the color-coded map in.
pop <- raster("pop_tot_stock_corine.tif")
lu <- raster("landcover_STK_clip.tif")
ind <- raster("industry_emission_stockholm_1113.tif")
trans <- raster("road_emission_stockholm_1113.tif")
res <- raster("resident_emission_stockholm_1113.tif")

crs(lu) <- crs(ind)
crs(trans) <- crs(ind)
crs(pop) <- crs(ind)
extent(lu) <- extent(ind)
extent(trans) <- extent(ind)
extent(pop) <- extent(ind)
lu <- resample(lu,ind,method = "bilinear")
trans <- resample(trans,ind,method = "bilinear")
res <- resample(res,ind,method = "bilinear")
pop <- resample(pop,ind,method = "bilinear")

# Select areas that need treatment
new <- raster(trans)
new[] <- 1

fun_1 <- function(x,y){ifelse( x == 10 | x == 11 , 2*y, y)} 
fun_2 <- function(x,y){ifelse( x > quantile(trans,0.90) , 3*y, y)}
fun_3 <- function(x,y){ifelse( x > quantile(ind,0.90) , 5*y, y)}
fun_4 <- function(x,y){ifelse( x > quantile(res,0.95) , 7*y, y)}
fun_5 <- function(x,y,z){ifelse( x > quantile(trans,0.90) & y > quantile(pop,0.75), 11*z, z)}
fun_6 <- function(x,y,z){ifelse( x >= 22 & y <= quantile(pop,0.8), 13*z, z)}
fun_7 <- function(x,y){ifelse( x > quantile(res,0.90) , 17*y, y)}


new <- overlay(lu,new,fun=fun_1)
new[is.na(new[])] <- 1
new <- overlay(trans,new,fun=fun_2)
new[is.na(new[])] <- 1
new <- overlay(ind,new,fun=fun_3)
new[is.na(new[])] <- 1
new <- overlay(res,new,fun=fun_4)
new[is.na(new[])] <- 1
new <- overlay(trans,pop,new,fun=fun_5)
new[is.na(new[])] <- 1
new <- overlay(lu,pop,new,fun=fun_6)
new[is.na(new[])] <- 1
new <- overlay(res,new,fun=fun_7)
new[is.na(new[])] <- 1

# add stockholm boundary
crop_extent <- readOGR("Kommuner_combined.shp")

crop_extent <- spTransform(crop_extent, CRS(proj4string(new)))

nbs <- mask(new, crop_extent)
nbs[is.na(nbs[])] <- 1
# writeRaster(r,'test.tif',options=c('TFW=YES'))
```

Start with one of the NbS approach implementation maps for the 50 cities in Europe.  
I used Stockholm as an example.  
```{r message=FALSE, warning=FALSE}
library(raster)
library(ggplot2)
library(tibble)
#library(rgdal)
nbs <- raster("Stockholm_emission/nbs.tif")
```

The stockholm map has a bunch of prime number representing potential NBS approaches.   
```{r}
#plot(nbs)
sort(unique(getValues(nbs)))
```

## Create distribution  
In the following table, each NbS approach may have direct "carbon reduction" effect, based on the "param" percentages referenced from the literature. E.g., the third line reads: road greening can reduce carbon emission from the transportation sector by 14%.   
Some NbS increases vegetation sequestration, based on the implementation areas on the map. Here the "param" is also values from the literature, while "adj" indicates how to convert/apply this parameter to the map. E.g. the second line reads: 1 km2 of urban forest can create carbon sequestration of 58.3-79.5 ton CO2 per year.    

|    Approach     |  ID |       Use         |    Sector      |   param   |       adj      |  var  |
|:----------------|-----|-------------------|----------------|-----------|----------------|-------|
|  Urban Forest   |  2  |  Carbon reduction |  all sectors   | 0.08%-2%  |           1    | uf_r  |
|  Urban Forest   |  2  |  Sequestration    |   per km2      | 58.3-79.5 |$30\times 30/10^6$| uf_s  |
|  Road Greening  | 11  |  Carbon reduction | transportation |    14%    |           1    | rd_r  |
|  Road Greening  |  3  |  Sequestration    |   per m2       | 0.0008304 |$30\times 30\times 0.3$| rd_s  |
|  Green belt     | 13  |  Sequestration    |   per m        |  0.07729  |      30        |  gb   |
|  Green roof     |  7  |  Carbon reduction |   res, ind     |  11-19%   |           1    | gr_r  |
|  Green roof     |  7  |  Sequestration    |   per m2       |1.703-1.889|$0.4\times 30\times 30/10^3$| gr_s  |
| GI for industry |  5  |  Carbon reduction |   industry     |  18.6%    |           1    | gind  |
| Green Access    | 17  |  Carbon reduction |  all sectors   |    10%    |           1    |  gc   |

This section determines the average, upper and lower boundaries of carbon reduction for each type of land use to create normal distributions.   
```{r}
set.seed(42)
###(2) Urban forest
# Carbon reduction
LO = 0.001
M = 0.01
UP = 0.022
S = (UP-LO)/6 # suppose UP-LO cover an area of 3*sigma
uf_r = rnorm(n = 1000, mean = M, sd = S)
# Sequestration
LO = 58.3
M = 79.5 
#UP = 216
S = (M-LO)/3 # suppose LO-M cover an area of 3*sigma
uf_s = rnorm(n = 1000, mean = M, sd = S)

###Road greening
#(11) Carbon reduction
LO = 0
M = 0.14 #14%
UP = 0.2
S = (UP-M)/3 # suppose M-LO cover an area of 3*sigma
rd_r = rnorm(n = 1000, mean = M, sd = S)
#(3) Sequestration
LO = 0
M = 0.00084 
S = (M-LO)/3 # suppose M-LO cover an area of 3*sigma
rd_s = rnorm(n = 1000, mean = M, sd = S)

###(13) Green belt
#Sequestration
LO = 0
M = 0.07729
S = (M-LO)/3 # suppose M-LO cover an area of 3*sigma
gb = rnorm(n = 1000, mean = M, sd = S)

###(7). Green roof 
# Sequestration
LO = 1.703
M = 1.78
UP = 1.889
S = (UP-LO)/6 # suppose UP-LO cover an area of 6*sigma
gr_s = rnorm(n = 1000, mean = M, sd = S)

# Carbon Reduction
LO = 0.11
UP = 0.19
M = 0.14
S = (UP-LO)/6 # suppose UP-LO cover an area of 3*sigma
gr_r = rnorm(n = 1000, mean = M, sd = S)

###(5). GI for ind (% of reduction of total INDUSTRY EMISSION)  
LO = 0.11
M = 0.186 
UP = 0.24
S = (UP-M)/3 # suppose UP-M cover an area of 3*sigma
gind = rnorm(n = 1000, mean = M, sd = S)

###(17). Improve green access 
# Carbon reduction
LO = 0.09
UP = 0.11
M = 0.103
S = (UP-LO)/6 
gc = rnorm(n = 1000, mean = M, sd = S)
```

## Sample a group of values.  
Run simulation once and get 9 values, corresponding to the table in the beginning.    
```{r}
df <- data.frame(c(2, 2, 11, 3, 13, 7, 7, 5, 17), 
                 c("carbon", "sequestration", "carbon", "sequestration","sequestration", "carbon", "sequestration", "carbon", "carbon"),
                 c(sample(uf_r, 1), sample(uf_s, 1), sample(rd_r, 1), sample(rd_s, 1), 
                   sample(gb, 1), sample(gr_r, 1), sample(gr_s, 1), sample(gind, 1), sample(gc, 1)))
colnames(df) <- c("id", "type", "param")
```

Add the "adj" column in df for calculation.   
```{r}
# Assumptions
CELL = 30  # Resolution/cell length
TREE = 0.3 # Tree shades cover 50% of streetscape cell areas
ROOF = 0.4 # Green roof can be implemented on 60% of cell areas

adj <-  c(1, CELL^2/10^6, # area adjusted from km2 to m2
          1, CELL^2*TREE, # area*tree shade percentage
          CELL, 1, ROOF*CELL^2/10^3, 1, 1)
df <- cbind.data.frame(df, adj)
rm(list=ls()[! ls() %in% c("df","nbs")])
```

### Build emission calculation functions
We know from the map of prime numbers (nbs), the areas for implementing each type of NBS (or NBS combination), and the prime number indicating specific approaches.   
First create a function to decompose the prime numbers:    

```{r}
# Prime Factorization  
prime_fac <- function(num) {
  up = floor(num**0.5 + 1)
  ans = c()
  for (i in 2:up){
    while (num %% i == 0){
      num = num/i
      ans = append(ans, i)
    }
  }
  if (num != 1){
    ans = append(ans, num)
  }
  return(ans)
}
```

As we have two approaches for emission reduction:  
* Based on area or existing physical features  
* Based on existing emission  
We will read in the emission maps for the subsequent calculation.   
```{r message=FALSE, warning=FALSE}
emission_ind <- raster("industry_emission_stockholm_0609.tif")
emission_res <- raster("resident_emission_stockholm_1113.tif")
emission_tran <- raster("road_emission_stockholm_1113.tif")
emission_res <- crop(emission_res, emission_tran)
emission_ind <- crop(emission_ind, emission_tran)

emission_res <- getValues(emission_res)
emission_ind <- getValues(emission_ind)
emission_tran <- getValues(emission_tran)
```

The following code block adds up carbon reduction and sequestration effects for each NbS approach.   
```{r}
# Set up variables  
res_reduction = 0
ind_reduction = 0
tran_reduction = 0
totalsequ = 0

# Create a new raster with the initial value 0 to record the accumulated spatial reductions
initial_raster <- raster(nrow = nbs@nrows, ncol=nbs@ncols)
values(initial_raster) = 0
# reductionmap <-  getValues(initial_raster)
sequestrationmap <- getValues(initial_raster)
source = getValues(nbs)

# for every prime number 
  # check if the prime number is factored by one of the 9 values in the table, 
   # if the prime number consists of 2 - calculate urban forest carbon reduction and sequestration
     ## carbon reduction is calculated by param (percentage) * emission (by sector)
     ## sequestration is calculated by area * param (carbon reduction per area unit) * adj (unit conversion)
   # if the prime number consists of  3 - calculate road green sequestration
# so on and so forth
#===Note: only calculated urban land (omitted value 2)==
for (num in unique(getValues(nbs))[-1]){ # remove number 1 - default number, not in the prime list
  # loop through each unique prime multiplication value in nbs,
  for (id in c(3, 7, 13)){ # sequestration for urban forest, road green, green belt and green roof
    if (id %in% prime_fac(num)){
      # if num contains a certain nbs value (e.g. 3), all cells that are equal to num should be implemented by road greening. Based on the parameters, sequestration will be param*adj on each cell
      PARAM = df[df$id == id & df$type == "sequestration", ]$param
      ADJ = df[df$id == id & df$type == "sequestration", ]$adj
      # draw map
      sequestrationmap[which(source %in% c(num))] = sequestrationmap[which(source %in% c(num))] + PARAM * ADJ 
      # accumulate emission reduction values
      totalsequ = totalsequ + sum(getValues(nbs) == num) * PARAM * ADJ
    }
  }
  #print(c(num, totalsequ))
  for (id in c(5, 7, 11, 17)){
    if (id %in% prime_fac(num)){
      PARAM = df[df$id == id & df$type == "carbon", ]$param
      if (id == 17){
        #reductionmap[which(source %in% c(num))] = reductionmap[which(source %in% c(num))] + emission_res[which(source %in% c(num))] * PARAM + emission_ind[which(source %in% c(num))] * PARAM+ emission_tran[which(source %in% c(num))] * PARAM
        res_reduction = res_reduction + sum(emission_res[which(source %in% c(num))], na.rm = TRUE) * PARAM
        ind_reduction = ind_reduction + sum(emission_ind[which(source %in% c(num))], na.rm = TRUE) * PARAM
        tran_reduction = tran_reduction + sum(emission_tran[which(source %in% c(num))], na.rm = TRUE) * PARAM
      }
      else if (id == 11){
        #reductionmap[which(source %in% c(num))] = reductionmap[which(source %in% c(num))] + emission_tran[which(source %in% c(num))] * PARAM * ADJ
        tran_reduction = tran_reduction + sum(emission_tran[which(source %in% c(num))], na.rm = TRUE) * PARAM
      }
      else if (id == 7){
        #reductionmap[which(source %in% c(num))] = reductionmap[which(source %in% c(num))] + emission_res[which(source %in% c(num))] * PARAM * ADJ + emission_ind[which(source %in% c(num))] * PARAM * ADJ
        res_reduction = res_reduction + sum(emission_res[which(source %in% c(num))], na.rm = TRUE) * PARAM
        ind_reduction = ind_reduction + sum(emission_ind[which(source %in% c(num))], na.rm = TRUE) * PARAM
      }
      else {
        #reductionmap[which(source %in% c(num))] = reductionmap[which(source %in% c(num))] + emission_ind[which(source %in% c(num))] * PARAM * ADJ
        ind_reduction = ind_reduction + sum(emission_ind[which(source %in% c(num))], na.rm = TRUE) * PARAM
      }
    }
  }
  #print(c(num, res_reduction, ind_reduction, tran_reduction, totalsequ))
}
```


## Produce maps
```{r}
# ===I only created sequestration maps in this section,
# ===if you also want to create carbon reduction maps by sector, un-comment the lines for making maps in the loop function above as well as the following three lines
# final_carbon <- raster(matrix(reductionmap, ncol = nbs@ncols, byrow = TRUE))
# extent(final_carbon) <- extent(nbs)
# crs(final_carbon) <- crs(nbs)
# 
final_sequestration <- raster(matrix(sequestrationmap, ncol = nbs@ncols, byrow = TRUE))
extent(final_sequestration) <- extent(nbs)
crs(final_sequestration) <- crs(nbs)
```

## Values and percentages  
How much do they take in the total emissions.   
```{r}
all <- unique(getValues(nbs))[-1]
# carbon reduction by sector
urban_res_emission <- sum(emission_res[which(source %in% c(all))], na.rm = TRUE) 
urban_ind_emission <- sum(emission_ind[which(source %in% c(all))], na.rm = TRUE) 
urban_tran_emission <- sum(emission_tran[which(source %in% c(all))], na.rm = TRUE) 
res_reduction/urban_res_emission
ind_reduction/urban_ind_emission
tran_reduction/urban_tran_emission
# sequestration
totalsequ/sum(emission_res+emission_ind+emission_tran, na.rm = TRUE) 
```
If you want to summarize results in a csv
```{r}
table <- tibble(
  Emission_Type = c("Residential", "Industry", "Transportation"),
  Total_Emission = c(urban_res_emission, urban_ind_emission, urban_tran_emission),
  Saving_Rate = c(paste0(round(res_reduction*100/urban_res_emission, 1), "%"),
                  paste0(round(ind_reduction*100/urban_ind_emission, 1), "%"),
                  paste0(round(tran_reduction*100/urban_tran_emission, 1), "%")),
  City_Name = "Stockholm")
table
#write.csv(table, "result.csv", row.names = FALSE)
```


## Make the waterfall graph   
```{r}
balance <- data.frame(desc = c('residential_emission', 'residential_reduction', 'industry_emission', 'industry_reduction', 'transportation_emission','transportation_reduction','sequestration'),
                      amount = c(urban_res_emission, -res_reduction, urban_ind_emission, -ind_reduction, urban_tran_emission, -tran_reduction, -totalsequ/10^3))
balance$desc <- as.factor(balance$desc)
levels(balance$desc) <- c("1","2","3","4","5","6","7")
balance$id <- seq_along(balance$amount)
balance$type <- ifelse(balance$amount>0, "emission", ifelse(balance$desc == "sequestration", "sequestration", "reduction"))

# end is the cumulative sum of the amount
# start is the end variable with a lag
balance$end <- cumsum(balance$amount)
balance$start <- c(0, head(balance$end, -1))
balance <- balance[, c(3,1,4,6,5,2)]

balance$type <- as.factor(balance$type)
ggplot(balance, aes(x=desc, fill=type))+
  geom_rect(aes(desc, xmin = id-0.45, xmax = id+0.45,
                ymin=end, ymax = start)) +
  theme_bw()+
  scale_x_discrete(labels=c("1" = "Residential", "2" = " ","3" = "Industry",
                            "4"= " ", "5" = "Transportation", "6" = " ", "7" = "Sequestration"),
                   name = "Emission and carbon reduction items by sector")+
  scale_y_continuous(name = "ton-CO2 /m²/yr")
  
```

